---
title: "Homework-26"
format: html
editor: visual
---

### Introduction

Practice running a one-way ANOVA by using the data on pea plant root lengths, measured in optical units for plants grown on different growth media.

Set things up:

```{r}
rm(list = ls())

library(ggfortify)
library(multcomp)
library(tidyverse)
library(here)
```

Import the dataset:

```{r}
peas <- read.csv(here("Data/s_r_pea_sections.csv"))
```

Change Treatment to a factor variable:

```{r}
peas$Treatment <- as.factor(peas$Treatment)
levels(peas$Treatment)
```

Part 1: Is there a difference in mean pea lengths?

Part 2: If there is a difference, for which growth media does the difference exist?

To answer these questions, follow the workflow introduced in Chapter 5.

### Plot the data

Build a histogram:

```{r}
ggplot(peas, aes(pea_section_length_ou))+
  geom_histogram()+
  facet_wrap(~Treatment, ncol = 1)
```

Build a box plot:

```{r}
ggplot(peas, aes(Treatment, pea_section_length_ou))+
  geom_boxplot()+
  coord_flip()
```

Estimate the results:

I assume that peas grown in the control medium will be significantly longer than other pea plants.

### Build the model


```{r}
pea_mod<- lm(pea_section_length_ou ~ Treatment, data = peas)
```

### Check model assumptions

```{r}
autoplot(pea_mod, smooth.colour = NA)
```

The distribution is fine. 

### Interpret results

```{r}
anova(pea_mod)
```

Since our p-value is small, there is a signficant effect of growth treatment on pea plant length.


```{r}
summary(pea_mod)
```

- Because 1% fructose 1% glucose is first alphabetically, the contrasts show us the difference in mean between the that and each other growth treatment.

- Control plants grew signficantly longer than the 1% glucose 1% fructose treatment, as did plants in the 2% sucrose treatment. Neither 2% fructose or 2% glucose were different from 1% fructose 1% glucose.


We are going to do "post hoc" tests of difference in means. To do so, we need the `multcomp` package, which we already run in the very beginning.

Conduct a Tukey's HSD test:

```{r}
phc1<- glht(pea_mod, linfct = mcp(Treatment = "Tukey"))

summary(phc1)
```

According to this result: 

- There is no difference in mean pea length for peas grown in 2% fructose vs. 1% f 1% s

- There is no difference in mean pea length for plants grown in 2% glucose vs. 1% f 1% s. 

- There is also no difference in mean pea length for peas grown in 2% glucose vs. those grown in 2% fructose. 

- All other groups were signficiantly different from one another.

We can also look at a "compact letter display" (cld) of these results:

```{r}
cld(phc1)
```

Or we can use the `emmeans()` function from the `emmeans` package

```{r}
library(emmeans)
ph2 <- emmeans(pea_mod, specs = "Treatment")
summary(ph2)
```

Then use the `pairs()` function to run the Tukey test

```{r}
pairs(ph2)
```

