---
title: "Homework 24"
format: html
editor: visual
---

## Overview

For your homework you will implement a data analysis using a dataset on crab growth measurements.

Set things up (hint: there is a new package to add with `library()`)

```{r}
rm(list = ls())
library(tidyverse)
library(ggplot2)
library(ggfortify) #evaluated model assumptions
library(here)
```

## The data

Let's pull in some data appropriate biologists: crab measurements. This dataset includes measurements of farmed crabs.

```{r}
crabs <- read.csv(here("DATA", "CrabAgePrediction.csv"))
glimpse(crabs)
```

One of the variables is categorical - let's fix that in the data.

```{r}
crabs$Sex <- as.factor(crabs$Sex)
```

## Question

Is there a relationship between height and weight in female crabs?

1.  Extract the data for just female crabs.

2.  Examine the response and predictor variables and use the flow chart from class to decide the appropriate statistical analysis (choose from: chi-square, 2-sample t-test or linear regression).

3.  Work through the data analysis **following all of the steps provided in chapter 5 of your book**. You may use class notes or homework keys as a guide.

## Plot the data

Extract the data for just female crabs:

```{r}
fcrabs <- crabs |> filter(Sex == "F")
```

Make a scatter plot:

```{r}
ggplot(fcrabs, aes(Height, Weight))+
     geom_point()+
     ylab("Weight of crab (g)")+
     theme_bw()
```

Looks like there's an outlier in height above 2. Remove it:

```{r}
fcrabs <- fcrabs |> filter(Height < 2)
ggplot(fcrabs, aes(Height, Weight))+
     geom_point()+
     labs(y = "Crab weight (g)", x = "Crab height (mm)")+
     theme_bw()

```

## Interpret the figure

There is a positive slope. We can also guesstimate the slope and intercept from the figure.

- The weight of the crabs ranges approximately between 2 and 76 grams

- And the crab heights varies approximately between 0.15 and 0.63 mm

Slope is rise/run, which in this case is about (76 - 2) = 74 for rise and 0.63 - 0.15 = 0.48 for run, so we get 74/0.48 = 154 as the slope.

Eyeballing the plot, we would also guess that the y-intercept is somewhere around -10

## Make a simple linear regression 

Use the `lm()` function:

```{r}
model_crabs <- lm(Weight ~ Height, data = fcrabs)
```

Ok, the model has been fit. We shuld consider assumptions before we look at the results: 

We need to use the `ggfortify` package and the `autoplot()` function within it:

```{r}
autoplot(model_crabs, smooth.color = NA)
```

Top right - Normal Q-Q plot: This evaluates the assumption that the residuals are normally distributed. We want the dots to fall pretty close to the line (observed = expected).

What does `smooth.color = NA` do? In the absence of this argument, the default presentation would be a wiggly line fitted by the regression. The NA suppresses that line.

## Interprete the results

Now that we know that our data meet the assumptions of a linear regression, we can look at and interpret the model.

We use two tools that we will use for every general (and generalized) model here on out: `anova()` and `summary()`.

`anova()` does not perform an ANOVA. Instead, it produces a classic anova table, the sums-of-squares table including the F-statistic, which is the ratio of variance explained by the explanatory variable to the leftover variance. As well, it produces an estimate of R\^2 and adjusted R\^2.

Use anova() and take a look:

```{r}
anova(model_crabs)
```

Look at the summary table:

```{r}
summary(model_crabs)
```

## From stats back to figure

Now let's make a figure that shows off our relationship in light of our statistical results.

```{r}
ggplot(fcrabs, aes(Height, Weight))+
     geom_point()+
     geom_smooth(method = "lm")+
     labs(x = "Crab height (mm)", y = "Crab weight (g)")+
     theme_bw()
```



