---
title: "Homework-25"
format: html
editor: visual
---

Clean R's environment.

```{r}
rm(list = ls())
```

## 5.6 Analysis of variance: the one-way ANOVA

Has one change in the variables: the explanatory variable is no longer continuous - it's either a factor or a categorical variable.

The dataset we will use focuses on Daphnia and their parasites.

### 5.6.1 Getting and Plotting the Data

There are two main questions:

1.  Do parasites alter growth rate?
2.  Does each of the three parasites reduce growth, compared with a control (no parasite treatment)?

Use the same workflow: Plot -\> Model -\> Check Assumptions -\> Interpret -\> Plot Again

```{r}
library(tidyverse)
library(ggfortify)
library(ggplot2)
library(here)
```

```{r}
daphnia <- read.csv(here("DATA", "Daphniagrowth.csv"))
```

```{r}
glimpse(daphnia)
```

This df has 3 variables: 'growth.rate' and 'parasite' are needed for a plot, 'rep' indicates the replication of each treatment.

Convert character variable into a factor.

```{r}
daphnia$parasite <- as.factor(daphnia$parasite)
```

Start by creating a box-and-whisker plot, which is effective for viewing variation in a response variable as a function of a grouping, categorical variable.

```{r}
ggplot(daphnia, aes(x = parasite, y = growth.rate)) +
  geom_boxplot() +
  theme_bw()
```

Although I was able to read the parasite names on my graph, there are instances where they don't fit quite right, so the chapter introduces us to coord_flip().

```{r}
ggplot(daphnia, aes(x = parasite, y = growth.rate)) +
  geom_boxplot() +
  theme_bw() +
  coord_flip()
#coord_flip() switches x- and y-axes (for view only)
```

There are a few things to take note of:

1.  There is a substancial variation in the daphnia growth rates among groups.
2.  Control treatment produces highest growth rate.
3.  There is likely to be a parasite treatment effect overall. The order of growth (low -\> high): P. ramosa \< M. bicuspidata \< P. perplexa

### 5.6.2 Construct the ANOVA

```{r}
model_grow <- lm(growth.rate ~ parasite, data = daphnia)
```

### 5.6.3 Check the assumptions

Plot four diagnostic graphs.

```{r}
autoplot(model_grow, smooth.colour = NA)
```

The graphs suggest that everything is just fine, and the pattern is whithin the expected bounds of variation found in samples of normal distribution.

### 5.6.4 Making an inference from a one-way ANOVA

We will interpret the output of the two tools for making inferences from linear models - anova() and summary ().

```{r}
anova(model_grow)
```

There is an evidence that the parasite treatment, comprising four levels of manipulation, has produced an effect.

-   For an ANOVA, null hypothesis means that all of the groups come from populations with the same mean.

-   F-value quantifies the ratio of the between-group variance to the within-group variance. Since the former is is large relative to the latter, this creates a large F-value.

-   A small p-value allows us to reject the null hypothesis.

### 5.6.5 Treatment contrasts

"Contrasts" - ways to produce information in a summary-like table. They are a way of expressing coefficients taken from statistical models.

Make a summary table:

```{r}
summary(model_grow)
```

-   R presents things in an alphabetical order.
-   "Control" group is not missing, but is labeled as an "Intercept".
-   In ANOVA framework, we can assume that the word 'Intercept' represents the first level of the alphabetically ordered treatment levels.
-   In a summary table, the numbers associated with each parasite are differences between growth rates associated with that parasite and the control. This is why they are all negative.

We can also get the means for each treatment level using group_by() and summarize(), and then calculate the contrast ourselves.

```{r}
#to get the mean growth rates:
sumDat <- daphnia %>%
  group_by(parasite) %>%
  summarize(meanGR = mean(growth.rate))
```

Now we can manually calculate how the control growth rate plus the treatment contrast for a given parasite in the summary table gives us the mean growth rate for that parasite.

E.g: 1.21 +(-0.73) = 0.48 which is a mean growth rate for P.ramosa

Finally, we can make a graph like in Figure 5.11. Specify two geom_points(): one for daphnia df aesthetics and another for sumDat df.

```{r}
ggplot(daphnia, aes(parasite, growth.rate, color = parasite))+
     geom_point(size = 1.5)+
     geom_point(data = sumDat, aes(parasite, meanGR, fill = parasite), size = 5)+
     coord_flip()+
     theme_bw()
```
