---
title: "Homework-16"
format: html
editor: visual
---
Clear R's environment and turn on "tidyverse" and "here" packages:

```{r}
rm(list = ls())
library(tidyverse)
library(here)
```

Create a new data frame called "cover":

```{r}
cover <- read.csv(here("DATA", "donnerville_percent_cover.csv"))
```


Examine the data set using at least two different functions. I chose glimpse(), str(), and summary()

```{r}
glimpse(cover) #gives quick overview of the data structure and content; shows columns horizontally
```
```{r}
str(cover) #more detailed view; useful for understanding data types
```
```{r}
summary(cover) #provides quick statistical overview of the data
```

Rename column headings:

```{r}
cover <- cover %>% rename(
  Plot_type = Plot.type,
  Plot_dim = Plot.dim,
  Cover_type = cover_type,
) #rename(new_name = old_name)
```

Convert "Site", "Plot_type" and "Cover_type" into factor variables:

```{r}
cover <- cover %>% mutate(
  Site = as.factor(Site),
  Plot_type = as.factor(Plot_type),
  Cover_type = as.factor(Cover_type)
)  #mutate() is used to modify columns, and as.factor() converted selected columns into factors
```

```{r}
glimpse(cover) #check if Site, Plot_type and Cover_type were converted from <chr> to <fct>
```

Convert "Date" into a date variable:

```{r}
library(lubridate) #checks and turn on "lubridate" package
```

```{r}
cover <- cover %>% mutate(Date = ymd(Date)) #lubridate can recognize common date formats like year-month-day
```

```{r}
glimpse(cover) #you can use str() or glimpse() to verify the changes or just open the data frame tab on top
```

Do some error checking on your data and fix any mistakes you find:

#Error fix in the "Site" column:
```{r}
fix1 <- which(cover$Site == "donnerville 1")
cover$Site[fix1] <- "Donnerville 1"
```

```{r}
fix2 <- which(cover$Site == "DONNERVILLE - 1")
cover$Site[fix2] <- "Donnerville 1"
```

```{r}
fix3 <- which(cover$Site == "donnerville 2")
cover$Site[fix3] <- "Donnerville 2"
```

```{r}
fix4 <- which(cover$Site == "donnerville-3")
cover$Site[fix4] <- "Donnerville 3"
```

#Error fix in the "Cover_type" column:
```{r}
fix5 <- which(cover$Cover_type == "bare")
cover$Cover_type[fix5] <- "Bare"
```
```{r}
fix6 <- which(cover$Cover_type == "vasc")
cover$Cover_type[fix6] <- "Vascular"   #for some reason it did not change one "vasc" into "Vascular" !!!
```

Determine the average percent cover by Cover_type (Bare, Bryophytes, Vascular plants) for Donnerville 1:
```{r}
cover %>% 
  group_by(Cover_type) %>% 
  summarise(meanPercent = mean(Percent))
```

Determine the average percent vascular plant cover across all sites:
```{r}
cover %>% 
  summarise(meanPercent = mean(Percent))
```

